---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidytuesdayR)
```


```{r, helper_functions}
count_NA_in_df = function(df){
  df %>% 
    map(~ sum(is.na(.x))) %>% 
    as_tibble()
}


var_summary_stats <- function(column, na.rm = TRUE) {
  # the argument column is expected to be numeric vector, as this function was designed 
  # to work with purrr::map (which passes dataframe columns as vectors to the function
  # being mapped)
  
  # return a single row dataframe containing the summary stats for column
  tibble(
    n = sum(!is.na(column)),
    min = min(column, na.rm = na.rm),
    q1 = quantile(column, probs = 0.25, na.rm = na.rm),
    median = median(column, na.rm = na.rm),
    q3 = quantile(column, probs = 0.75, na.rm = na.rm),
    max = max(column, na.rm = na.rm),
    iqr = IQR(column, na.rm = na.rm),
    mean = mean(column, na.rm = na.rm),
    sd = sd(column, na.rm = na.rm)
  )
}

df_summary_stats <- function(df){
  
  # get the names of the numeric variables in the dataframe
  variable_name <- df %>%
    select_if(is.numeric) %>% 
    names()

  # return a dataframe containing summary stats for each numeric variable in df
  df %>%
    select_if(is.numeric) %>% 
    purrr::map_dfr(var_summary_stats) %>% 
    add_column(variable_name, .before = 1)
}

`%nin%` = Negate(`%in%`)
```


```{r}

# Argument for parsing commisioning date as a double, which throw parsing errors - 
# col_types = cols(commissioning_date = col_double()) 
wind_turbine <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-10-27/wind-turbine.csv')
wind_turbine


```

## Tidy data

check types - `commissioning_date` as character, all other types look ok at first glance.

Errors when try to coerce `commissioning_date` to a numeric, as the column includes up to three commissioning dates (e.g. 2005/2006/2012)

```{r}
wind_turbine %>% 
  distinct(commissioning_date)
```
So, separate the strings for commissioned data into upto three variables (`com_date_1`, `com_date_2`, `com_date_3`) and then coerce these from characters to numerics 

```{r}
wind_turbine <- wind_turbine %>% 
  separate(commissioning_date, sep = "/", into = c("com_date_1", "com_date_2", "com_date_3")) %>%
  mutate(com_date_1 = as.numeric(com_date_1),
         com_date_2 = as.numeric(com_date_2),
         com_date_3 = as.numeric(com_date_3), 
         most_recent_com_date = pmax(com_date_1, com_date_2, com_date_3, na.rm = TRUE)
         )

wind_turbine
```
Hoping to get some interesting information about lifespans of turbines (where multiple commissioning dates where provided in the dataset), but little interesting or useful information in here.
```{r}

wind_turbine %>% 
  # calculate lifespans where multiple commissioning dates are provided
  mutate(lifespan_1 = com_date_2 - com_date_1,
         lifespan_2 = com_date_3 - com_date_2) %>% 
  select(lifespan_1, lifespan_2) %>% 
  df_summary_stats()
```


`turbine_number_in_project` - looks like to variable combined into a single column (turbine number / number of turbines in project)

```{r}
wind_turbine <- wind_turbine %>% 
  separate(turbine_number_in_project, sep = "/", into = c("turbine_number", "number_of_turbines_in_project")) %>% 
  mutate(turbine_number = as.numeric(turbine_number),
         number_of_turbines_in_project = as.numeric(number_of_turbines_in_project))
wind_turbine

```
A quick check on the number of NAs in the wind turbines dataframe. Nothing too alarming. `turbine_rated_capacity_k_w` has 220 missing entries, so keep that in mind if using it in the analysis later on.

```{r}
count_NA_in_df(wind_turbine)
```



```{r}
plot_core <- ggplot(data = wind_turbine,
                    mapping = aes(x = most_recent_com_date)
                    )

plot_core + geom_histogram() +
  facet_wrap(~manufacturer)
```

Proportion of turbines commissioned by year- top 5 manufacters + others

```{r}

wind_turbine

# subset data for interest
wind_turbine_focus <- wind_turbine %>% 
  select(turbine_rated_capacity_k_w, manufacturer, most_recent_com_date)

# identify top 5 manufacturers (by total number of wind turbines)
manufacturers <- wind_turbine_focus %>% 
  group_by(manufacturer) %>% 
  summarise(num_turbines = n()) %>%
  mutate(manu_ranking = min_rank(desc(num_turbines))) %>% 
  arrange(desc(num_turbines))

num_top_manu <- 4
top_manu <- as_vector(manufacturers[1:num_top_manu, 'manufacturer'])

# group all manufacturers outside top 5 together
wind_turbine_focus <- wind_turbine_focus %>% 
  mutate(manufacturer = replace(manufacturer, manufacturer %nin% top_manu, "Others"))

# add ranking for each manufacturer (based on total number of wind turbines)
wind_turbine_focus <- wind_turbine_focus %>% 
  left_join(manufacturers) %>% 
  select(-num_turbines) %>% 
  mutate(manu_ranking = replace(manu_ranking, manufacturer == "Others", 6))

# for each year calculate the amount of capacity and number of turbines added 
wind_turbine_focus <- wind_turbine_focus %>% 
  group_by(most_recent_com_date) %>% 
  mutate(annual_capacity_added = sum(turbine_rated_capacity_k_w, na.rm = TRUE),
         annual_turbines_added = n()
         ) %>% 
  ungroup()

wind_turbine_focus

# for each year calculate the proportion of turbines supplied and capacity added 
# by each manufacturer
wind_turbine_focus <-  wind_turbine_focus %>% 
  group_by(most_recent_com_date, manufacturer) %>%
  summarise(capacity_added = sum(turbine_rated_capacity_k_w, na.rm = TRUE),
            turbines_added = n(),
            prop_capacity_added = capacity_added / annual_capacity_added,
            prop_turbines_added = turbines_added / annual_turbines_added,
            annual_capacity_added = annual_capacity_added,
            annual_turbines_added = annual_turbines_added,
            manu_ranking = manu_ranking
            ) %>% 
  distinct() %>% 
  ungroup()

# confirm proportions sum to approx. 1 for each year
wind_turbine_focus %>% 
  group_by(most_recent_com_date) %>% 
  summarise(sum_prop_cap = sum(prop_capacity_added, na.rm = TRUE),
            sum_prop_turb = sum(prop_turbines_added, na.rm = TRUE))

count_NA_in_df(wind_turbine_focus)

# plot turbines added over time
p <- ggplot(data = wind_turbine_focus,
                 mapping = aes(x = most_recent_com_date,
                               y = annual_turbines_added)
                 )

p + geom_col()

# plot capacity added over time
p <- ggplot(data = wind_turbine_focus,
                 mapping = aes(x = most_recent_com_date,
                               y = annual_capacity_added)
                 )

p + geom_col()

# plot capacity added over time
p <- ggplot(data = wind_turbine_focus,
                 mapping = aes(x = most_recent_com_date,
                               y = annual_capacity_added)
                 )

p + geom_col()




```
```{r, fig.height=7}

min_year <- min(wind_turbine_focus$most_recent_com_date)
max_year <- max(wind_turbine_focus$most_recent_com_date)

x_limits = c(min_year, max_year)

# plot proportion of annual capacity added for each manufacturer over time
p_market_share <- ggplot(data = wind_turbine_focus,
                 mapping = aes(x = most_recent_com_date,
                               y = prop_capacity_added,
                               fill = reorder(manufacturer, manu_ranking))
                 )

p_market_share <- p_market_share + geom_col() +
  
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set3") +
  
  labs(x = "",
       y = "Market share",
       fill = "Manufacturer",
       title = "Shares of the Growing Canadian Wind Power Market by the Big Four Manufacturers",
       subtitle = "Data source: open.canada.ca") +
  
  theme_minimal()


# plot market growth
p_market_growth <- ggplot(data = wind_turbine_focus,
                          mapping = aes(x = most_recent_com_date,
                                        y = annual_capacity_added)
                 )
p_market_growth <- p_market_growth + 
  
  geom_line(colour = "grey70", size = 1.5) +
  
  scale_x_continuous(limits = x_limits) +
  
  labs(x = "",
       y = paste("Total capacity commissioned (megawatts)"),
       caption = "Market share: the proportion of the total wind energy generation
       capacity commissioned in a given year") +
  
  theme_minimal()


  # +
  #facet_wrap(~ reorder(manufacturer, manu_ranking), ncol = 2)

# p + geom_line() +
#    facet_wrap(~ reorder(manufacturer, manu_ranking), ncol = 2) +
p_market <- cowplot::plot_grid(p_market_share, p_market_growth,
                     ncol = 1, rel_heights = c(3, 2),
                     align = 'v', axis = 'lr')

p_market

```

```{r}

wind_turbine_focus %>% 
  mutate(bin = cut_width(most_recent_com_date, width = 5))
  

# plot proportion of annual capacity added for each manufacturer over time
p_market_share_2 <- ggplot(data = wind_turbine_focus,
                           mapping = aes(x = most_recent_com_date,
                                         y = prop_capacity_added,
                                         colour = reorder(manufacturer, manu_ranking)
                                         )
                           )


p_market_share_2 + geom_point() +
  
  scale_x_continuous(limits = x_limits) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set3") +

  labs(x = "",
       y = "Market share",
       fill = "Manufacturer",
       title = "Share of the Canadian Wind Power Market by the Big Four Manufacturers",
       subtitle = "Data source: open.canada.ca") +

  theme_minimal()


```


```{r}
plot_core <- ggplot(data = wind_turbine,
                    mapping = aes(x = most_recent_com_date,
                                  y = turbine_rated_capacity_k_w)
                    )

plot_core + geom_jitter(alpha = 0.4) + geom_smooth()
```

Could look at some average wind farm size by year. But not as straight forward as I first thought. The below is average size of wind farms where wind turbines have been commissioned in the year in question.

```{r}

wind_turbine %>% 
  group_by(most_recent_com_date) %>% 
  summarise(mean(number_of_turbines_in_project, na.rm = TRUE))
```
could look at capacity added per year

```{r}
wind_turbine

wind_turbine %>% 
  select()


```

